Spin canonical divisors of log stacky curves

### 2015-06-25 stacky_write_up/spin_cring.tex
remove redundancy from admiss def
edited genus 1 induction

\begin{thm}
\label{thm:ram_order_ind}
Suppose $(\sx', 0, \halfcan')$ is a log spin curve with coarse
space $X'$ and signature $(g; e_1', \ldots, e_r'; \delta)$ and
$(\sx', J)$ is admissible with generators $x_j, y_{i, e_i'} \in R' =
R_{L'}$ with $i \in J$ as in Definition ~\ref{defn:admissible}. Let
$(\sx, 0, \halfcan)$ be another log spin curve
with coarse space $X = X'$ and signature $(g; e_1, \ldots, e_r;
\delta)$ such that $e_i = e_i' + 2$ for all $i \in J$ and
$e_j = e_j'$ for $j \notin J$. Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i, e_i} \in
		H^0(\sx, e_i(K_\sx))$ so that
		\begin{align*}
			-ord_{Q_i}(y_{i, e_i}) = \frac{e_i - 1}{2}
		\end{align*}
		and
		\begin{align*}
			\frac{-\ord_{Q_j}(y_{i, e_i})}{\deg (y_{i, e_i})} \leq 
\subhalf{
			e_j'} - \frac{1}{\deg(y_{i, e_i})e_j'}
		\end{align*}
		for all $j \in J$ with $j \neq i$.
	\item[(b)] The elements $y_{i, e_i'}^a y_{i, e_i}^b$ with $a \geq 
0,
		b > 0$ span $R$ over $R'$ and the elements $y_{i, e_i}$ minimally
		generate $R$ over $R'$.
	\item[(c)] Equip $k[y] = k[y_{i, e_i}]_{i \in J}$ and $k[x]$ with
		any graded monomial order and $k[y, x] = k[y] \otimes k[x]$ with
		block order. Let $R = k[y, x]/I.$ Then,
		\begin{align*}
			in_\prec(I) = in_\prec(I')k[x, y] + \langle y_{i, e_i}z \mid z 
			\neq y_{i, e_i}, y_{i, e_i'} \rangle
		\end{align*}
		as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by
		products of two monomials and that for all $i \in J$, we have $e
		_i > \deg z$ for any generator $z$ of $R'.$ Then any set of
		minimal generators for $I'$ together with any set of relations 
as in 
		part (c) minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}

@@

Finally, if $(\sx, J)$ is admissible then so is $(\sx,W)$ for any
$W \subset J$ by Lemma ~\ref{lem:admissible_subset}.
\end{proof}

@@

\begin{prop}
\label{prop:relations_generation}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K + \Delta$. Then $I_L$ is generated in degree at most $11$ if $\Delta = 0$ and degree at most $10$ if $\Delta > 0$.
\end{prop}
\begin{proof}
Suppose $f \in I_L$ with $\deg f \geq 12$. To complete the proof, it suffices to show $f \in I_{L,d}'$. By Lemma ~\ref{lem:reducing_degree}, it suffices to show $'s_1g+s_2h \in I_{L,d}'$. 
Let $V \subset H^0(K)$ be the vector subspace generated by $\phi(s_1)$ and $\phi(s_2)$, and consider the map
$$\begin{tikzcd}
V \otimes H^0((\deg f - 2)L) \ar {r}{f} & H^0((\deg f)L),
\end{tikzcd}$$
we know that $\phi(s_1)\phi(g) + \phi(s_2) \phi(h) \mapsto 0.$
So, using the refined version of the basepoint-free pencil trick, as shown in the proof of \cite{saint-donat:proj}[Lemma 2.6], if $\phi(s_1) g + \phi(s_2)h \mapsto 0,$ there exists some $\rho \in k[x_1, \ldots, x_n]$ so that $\phi(\rho) \in H^0((\deg f - 4)L)$ satisfies $\phi(g) = \phi(s_2)\phi(\rho)$ and $\phi(h) = -\phi(s_1)\phi(\rho).$ Therefore, $g \equiv s_2 \rho \bmod I_{k,d-2}$ and $h \equiv -s_1 \rho \bmod I_{k,d-2}$. Therefore,
\begin{align*}
	s_1g + s_2h \equiv s_1s_2\rho + s_2(-s_1 \rho) \equiv 0 \bmod J_{L,d}.
\end{align*}
\end{proof}

\begin{rem}
\label{rem:relations_generation_ten}
In \cite{neves:halfcan}[Theorem III.19], Neves shows that if $L=P$ is a half canonical divisor on a genus 2 curve, then $R_L$ has a relation in degree 10. Therefore, the bound from Proposition ~\ref{prop:relations_generation}, that $I_L$ is generated in degree at most 11 is close to sharp.
\end{rem}

@@

@@

\todo{Aaron: It becomes a bit tedious for this proof to alternatingly cite ~\ref{lem:admissible_subset} and ~\ref{thm:ram_order_ind} is there a reason these were separated/can we merge the back together? Better yet, this should probably be a corollary somewhere, since we have to use this implicitly in all three genus cases.}
Second, we deal with the case that $\sx$ has at least two stacky points. If $(\sx',\Delta,L')$ is a spin canonical curve so that $L' = P - Q + \frac{1}{3}P_1+ \frac{1}{3}P_2$, then, by Example ~\ref{eg:base- 1-33}, it follows that $(\sx', \{3,3\})$ is admissible and satisfies the hypotheses of Theorem ~\ref{thm:ram_order_ind}. Therefore, applying Lemma ~\ref{lem:n-3s-admissible} $r-2$ times, we see that for $(\sx',\Delta',L')$ with $L' = P - Q + \sum_{i = 1}^{r}\frac{1}{3}P_i,$ the pair $(\sx',\{3, \ldots, 3\})$ is admissible. Assume for convenience that $e_1 \leq e_2 \leq \cdots \leq e_r$. Applying Theorem ~\ref{thm:ram_order_ind} $\frac{e_1-3}{2}$ times, and letting $J_1 = \{e_1, \ldots, e_1\}$ so that $\# J_1 = r$, we obtain that the theorem holds for $(\sx',\Delta',L')$ with $L' = P - Q + \sum_{i = 1}^{r}\subhalf{e_1}P_i,$ and that $(\sx',J_1)$ is admissible. Hence, by Lemma ~\ref{lem:admissible_subset}, it follows that $(\sx',J')$ is admissible, with $J' \subset \{e_1, \ldots, e_1\},$ with $\# J' = r- 1$. Then, applying Theorem ~\ref{thm:ram_order_ind} $\frac{e_2-e_1}{2}$ times, and letting $J_2 = \{e_2, \ldots, e_2\}$ so that $\# J_2 = r- 1$, we obtain that the theorem holds for $(\sx',\Delta',L')$ with $L' = P - Q + \subhalf{e_1}P_1 + \sum_{i =2}^{r}\subhalf{e_2}P_i,$ and that $(\sx',J_1)$ is admissible. Continuing to altenatively apply Lemma ~\ref{lem:admissible_subset} and Theorem ~\ref{thm:ram_order_ind}, we eventually obtain the theorem holds for $L' = P - Q + \sum_{i = 1}^{r}\subhalf {e_i}P_i$, as desired.

@@

\section{Further Research}
\label{sec:further-questions}
\begin{itemize}
	\item More on fractional modular forms
	\item Make explicit generators and relations for base case of non-stacky higher genus half-canonical divisors
	\item In the case $g \geq 2$, can the bound from Proposition ~\ref{prop:relations_generation} be reduced from 11 to 10 (which would then be sharp by Remark ~\ref{rem:relations_generation_ten})? \todo{Possibly think more about this} 
	\item Generic Initial ideal (and other genericness properties?)
\end{itemize}

>>>>

\begin{thm}
\label{thm:ram_order_ind}
Suppose $(\sx', \Delta, \halfcan')$ \todo{For some reason this $\Delta$ was set to 0. Was there any reason for this?} is a log spin curve with coarse
space $X'$ and signature $(g; e_1', \ldots, e_r'; \delta)$ and
$(\sx', J)$ is admissible with generators $x_j, y_{i, e_i'} \in R' =
R_{L'}$ with $i \in J$ as in Definition ~\ref{defn:admissible}. Let
$(\sx, \Delta, \halfcan)$ be another log spin curve
with coarse space $X = X'$ and signature $(g; e_1, \ldots, e_r;
\delta)$ such that $e_i = e_i' + 2$ for all $i \in J$ and
$e_j = e_j'$ for $j \notin J$. Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i, e_i} \in
		H^0(\sx, e_i(K_\sx))$ so that
		\begin{align*}
			-ord_{Q_i}(y_{i, e_i}) = \frac{e_i - 1}{2}
		\end{align*}
		and
		\begin{align*}
			\frac{-\ord_{Q_j}(y_{i, e_i})}{\deg (y_{i, e_i})} \leq 
\subhalf{
			e_j'} - \frac{1}{\deg(y_{i, e_i})e_j'}
		\end{align*}
		for all $j \in J$ with $j \neq i$.
	\item[(b)] The elements $y_{i, e_i'}^a y_{i, e_i}^b$ with $a \geq 
0,
		b > 0$ span $R$ over $R'$ and the elements $y_{i, e_i}$ minimally
		generate $R$ over $R'$.
	\item[(c)] Equip $k[y] = k[y_{i, e_i}]_{i \in J}$ and $k[x]$ with
		any graded monomial order and $k[y, x] = k[y] \otimes k[x]$ with
		block order. Let $R = k[y, x]/I.$ Then,
		\begin{align*}
			in_\prec(I) = in_\prec(I')k[x, y] + \langle y_{i, e_i}z \mid z 
			\neq y_{i, e_i}, y_{i, e_i'} \rangle
		\end{align*}
		as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by
		products of two monomials and that for all $i \in J$, we have $e
		_i > \deg z$ for any generator $z$ of $R'.$ Then any set of
		minimal generators for $I'$ together with any set of relations 
as in 
		part (c) minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}

@@

Finally, if $(\sx , J)$ is admissible then so is $(\sx,W)$ for any
$W \subset J$ by Lemma ~\ref{lem:admissible_subset}.
\end{proof}

\begin{cor}
\label{cor:raise-stacky-order}
Suppose $(\sx',L',J')$ is admissible with signature $\sigma' = (e_1',\ldots,e_r')$ with $J' = \{t,t+1,\ldots,r\}$ and $e_1' \leq e_2' \leq \cdots \leq e_t' = e_{t+1}' =\cdots = e_r'$
so that $(\sx',\Delta, L)$ satisfies the conditions of Theorem ~\ref{thm:ram_order_ind} and Theorem ~\ref{thm:general_generators_relations}. Then, for any spin curve $(\sx,\Delta,L)$ so that $\sx$ and $\sx'$ have the same coarse space $X = X'$ with the same set of ramified points, and $\sx$ has signature $(g;e_1,\ldots, e_r;\delta)$ with $e_1 \leq e_2 \leq \cdots e_r$ so that
$$
e_i \begin{cases}
	= e_i' &\text{ if }i \notin J\\
	\leq e_i' &\text{ if } i \in J
\end{cases}$$
then Theorem ~\ref{thm:general_generators_relations} holds for $(\sx,\Delta,L)$
\end{cor}
\begin{proof}
For $t \leq i \leq r$. Let $(\sx_i,\Delta, L_i)$ be the log stacky spin curve with the same coarse space as $L'$ and with the same stacky points as $(\sx',\Delta,L),$ having signature $(g;e_1, \ldots e_{i-1}, e_i, e_i, \ldots, e_i;\delta),$ and let 
$J_i = \{i,\ldots, r\}.$ Note that $(\sx_k,L_k,J_0) = (\sx',L',J')$ and $(\sx_r ,L_r ,J_r) = (\sx, L, J)$.

Let $(*_i)$ denote the condition that $(\sx_i,\Delta, L_i)$ satisfies the conditions $(\sx',\Delta, L)$ satisfies the conditions of Theorem ~\ref{thm:ram_order_ind} and Theorem ~\ref{thm:general_generators_relations} and $(\sx_i, L_i, J_i)$ is admissible.

Since $(*_0)$ holds by assumption, by induction on $i$, it suffices to show that if $(*_i)$ holds then so does $(*_{i+1}).$ 
Indeed, by applying Lemma ~\ref{lem:admissible_subset}, it follows $(\sx_i, L_i, J_{i+1})$ is admissible, as $J_{i+1} \subset J_i$. Then, applying Theorem ~\ref{thm:ram_order_ind} $\frac{e_{i+1}-e_i}{2}$ times, we see $(*_{i+1})$ holds.
\end{proof}

@@

\begin{prop}
\label{prop:relation_11}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K + \Delta$. Then $I_L$ is generated in degree at most $11$ if $\Delta = 0$ and degree at most $10$ if $\Delta > 0$.
\end{prop}
\begin{proof}
Suppose $f \in I_L$ with $\deg f \geq 12$. To complete the proof, it suffices to show $f \in I_{L,d}'$. By Lemma ~\ref{lem:reducing_degree}, it suffices to show $'s_1g+s_2h \in I_{L,d}'$. 
Let $V \subset H^0(K)$ be the vector subspace generated by $\phi(s_1)$ and $\phi(s_2)$, and consider the map
$$\begin{tikzcd}
V \otimes H^0((\deg f - 2)L) \ar {r}{f} & H^0((\deg f)L),
\end{tikzcd}$$
we know that $\phi(s_1)\phi(g) + \phi(s_2) \phi(h) \mapsto 0.$
So, using the refined version of the basepoint-free pencil trick, as shown in the proof of \cite{saint-donat:proj}[Lemma 2.6], if $\phi(s_1) g + \phi(s_2)h \mapsto 0,$ there exists some $\rho \in k[x_1, \ldots, x_n]$ so that $\phi(\rho) \in H^0((\deg f - 4)L)$ satisfies $\phi(g) = \phi(s_2)\phi(\rho)$ and $\phi(h) = -\phi(s_1)\phi(\rho).$ Therefore, $g \equiv s_2 \rho \bmod I_{k,d-2}$ and $h \equiv -s_1 \rho \bmod I_{k,d-2}$. Therefore,
\begin{align*}
	s_1g + s_2h \equiv s_1s_2\rho + s_2(-s_1 \rho) \equiv 0 \bmod J_{L,d}.
\end{align*}
\end{proof}

\begin{rem}
\label{rem:relations_generation_ten}
In \cite{neves:halfcan}[Theorem III.19], Neves shows that if $L=P$ is a half canonical divisor on a genus 2 curve, then $R_L$ has a relation in degree 10. Therefore, the bound from Proposition ~\ref{prop:relation_11}, that $I_L$ is generated in degree at most 11 is close to sharp.
\end{rem}

@@

\begin{thm}
\label{thm:general_generators_relations}
Let $g \geq 2$ and let $(\sx, \Delta, \halfcan)$ log spin curve over a perfect field $k$, so
that $\sx$ has signature $\sigma = (g; e_1, \ldots, e_r; \delta)$. Then the
canonical ring

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r).$ If $\Delta = 0$ then 
$R(\sx,\Delta, \halfcan)$ has minimal relations in degree at most $\max(11, 2e_1,
\ldots, 2e_r).$ Otherwise, $R(\sx, \Delta, \halfcan)$ has minimal relations in degree at most
$\max(10, 2e_1, \ldots, 2e_r).$
\end{thm}

\begin{proof}
By Lemma ~\ref{lem:generation_5} and Proposition ~\ref{prop:relation_11}, we are done in the case that there are no stacky points. 
\end{proof}

@@

Second, we deal with the case that $\sx$ has at least two stacky points. If $(\sx',\Delta,L')$ is a spin canonical curve so that $L' = P - Q + \frac{1}{3}P_1+ \frac{1}{3}P_2$, then, by Example ~\ref{eg:base- 1-33}, it follows that $(\sx', \{3,3\})$ is admissible and satisfies the hypotheses of Theorem ~\ref{thm:ram_order_ind}. Therefore, applying Lemma ~\ref{lem:n-3s-admissible} $r-2$ times, we see that for $(\sx',\Delta',L')$ with $L' = P - Q + \sum_{i = 1}^{r}\frac{1}{3}P_i,$ the pair $(\sx',\{3, \ldots, 3\})$ is admissible. Assume for convenience that $e_1 \leq e_2 \leq \cdots \leq e_r$. Then, by Corollary ~\ref{cor:raise-stacky-order}, this theorem holds for $L' = P - Q + \sum_{i = 1}^{r}\subhalf {e_i}P_i$, as desired.

@@

\section{Further Research}
\label{sec:further-questions}
\begin{itemize}
	\item More on fractional modular forms
	\item Make explicit generators and relations for base case of non-stacky higher genus half-canonical divisors
	\item In the case $g \geq 2$, can the bound from Proposition ~\ref{prop:relation_11} be reduced from 11 to 10 (which would then be sharp by Remark ~\ref{rem:relations_generation_ten})? \todo{Possibly think more about this} 
	\item Generic Initial ideal (and other genericness properties?)
\end{itemize}
